---
title: "BST 270 Individual Project"
author: "Roman Yan"
output: html_notebook
---

This file documents an attempt to reproduce the first two tables in the article [__A Better Way To Evaluate NBA Defense__](https://fivethirtyeight.com/features/a-better-way-to-evaluate-nba-defense/).

Firstly, load the packages and data. The packages used in this document are `dplyr v1.0.5` and `DT v0.31`.  The codes in the document are developed in `R 4.0.2`. The [dataset](https://raw.githubusercontent.com/fivethirtyeight/data/master/nba-draymond/draymond.csv) contains the DRAYMOND ratings (Defensive Rating Accounting for Yielding Minimal Openness by Nearest Defender) for every player since the 2013-14 season.

```{r}
library(dplyr)
data <- read.csv("draymond.csv")
```


### The First Table: The best defender according to DRAYMOND is Draymond

_NBA players by DRAYMOND* defensive ratings, based on opponents’ shooting data in the regular season and playoffs, with a minimum of 10,000 possessions played since 2013-14_

For each player, the possessions is their sum of all possessions across years in the dataset, and the DRAYMOND point is their average DRAYMOND points across years weighted by the yearly possession. The table is sorted by the average DRAYMOND points in descending order.

```{r}
data %>% group_by(PLAYER = player) %>% 
  mutate(weight = possessions/sum(possessions)) %>% # create weights using proportions of possessions
  mutate(weighted_DRAYMOND = DRAYMOND*weight) %>% # calculate the weighted DRAYMOND point 
  summarise(`POSSESSIONS PLAYED` = round(sum(possessions)), `DRAYMOND RATING` = round(sum(weighted_DRAYMOND),2)) %>% # round the displayed numbers
  filter(`POSSESSIONS PLAYED` >= 10000) %>% # only display for players with greater than 10,000 possessions
  arrange(desc(`DRAYMOND RATING`)) %>% # sort the rating in descending order
  mutate_at(vars(`DRAYMOND RATING`), ~if_else(. > 0, paste0("+", .), paste0(.))) %>% # add the "+" sign for positive DRAYMOND ratings 
  DT::datatable() # render interative table ouput
```

  *Defensive Rating Accounting for Yielding Minimal Openness by Nearest Defender

### The Second Table: The best DRAYMOND defenders of 2018-19

_NBA players by DRAYMOND* defensive ratings, based on opponents’ shooting data in the regular season and playoffs, with a minimum of 2,000 possessions played in 2018-19_

The results can be obtained by filtering the original dataset on the season and the minimum possessions as the. The table is sorted by the average DRAYMOND points in descending order.

```{r}
data %>% filter(season == 2019, possessions >= 2000) %>% # extract data for 2018-19 season and players with greater than 2,000 possessions in the season
  select(PLAYER = player, `POSSESSIONS PLAYED` = possessions, `DRAYMOND RATING` = DRAYMOND) %>% # rename columns to match the output in the article
  arrange(desc(`DRAYMOND RATING`)) %>% # sort the DRAYMOND rating in descending order
  mutate_at(vars(`DRAYMOND RATING`), ~round(., 2)) %>% mutate_at(vars(`POSSESSIONS PLAYED`), ~round(.)) %>% # round the numbers
  mutate_at(vars(`DRAYMOND RATING`), ~if_else(. > 0, paste0("+", .), paste0(.))) %>%  # add the "+" sign for positive DRAYMOND ratings 
  DT::datatable() # render interative table ouput
```

 *Defensive Rating Accounting for Yielding Minimal Openness by Nearest Defender
